Prefix-free graphs and suffix array construction in sublinear space

### 2023-06-24 manuscript/conclusion.tex
added conclusion

\section{Conclusion}
\subsection{Application}
There are several proposed data structures implicitly relying on the structure
of prefix-free graphs.

Rpair: rescaling repair with rsync \cite{2019gagie}
Prefix-free parsing for building big BWTs \cite{2019boucher}
MONI and r-index \cite{rossi2022moni,gagie2020fully}
Prefix-free Wheeler graphs \cite{2022pfwg}
MARIA \cite{2022maria}

All of these works implement their own version of prefix-free graphs.
We argue that separating prefix-free graphs as a standalone data structure can
bring several benefits:

\begin{itemize}
    \item reduce the complexity of the presentation
    \item allowing for better optimization of algorithms related to prefix-free graphs
    \item supporting the theoretical research by clearly deliminating the relevant terms
    \item allowing to bring prefix-free graphs closer to the biological data
\end{itemize}

Related to the last point, we remark that a possible future research can be in
the proper choice of trigger words.
As an example, we used stop codons in our experiments.
This could create prefix-free graphs capturing some biologically relevant phenomena.
Other options are recombination hotspots, highly repetitive elements, different
binding sites and sites with increased breakage.

>>>>

\section{Conclusion}
We introduced the prefix-free graph as a standalone data structure in pangenomics and showed how to use it to build a suffix array of a pangenome in sublinear space.
There are several possible directions for future research on prefix-free graphs.
The first direction may be to answer the question of which additional stringological algorithms and data structures can use prefix-free graphs to improve their applicability to such vast and repetitive datasets as pangenomic data.
Next, the choice of trigger words is, so far, mostly undiscovered territory.
From a computational perspective, the set of trigger words minimizing the size of a resulting prefix-free graph is desirable since it would allow the analysis of larger datasets.
Furthermore, trigger words offer great flexibility in the construction process.
We hypothesize this flexibility could be used for the construction of prefix-free graphs based on biologically significant strings.
In our experiments, we used the stop codon sequences as the trigger words, but other motives, such as different binding sites, recombination hotspots, or repetitive elements, are possible and could better capture some biological phenomena.
